IRT procedures for shortening tests

### 2024-10-10 the-good-the-bad-the-ugly.qmd
non funziona niente

---
title: "IRT procedures for shortening tests"
author: "Ottavia M. Epifania & Friends"
format: 
  revealjs: 
    theme: mytheme.scss
background-opacity: "0.45"
title-slide-attributes: 
   data-background-image: "https://psicostat.github.io/shine-bright/gbu.jpg"
   data-background-size: 350px, cover
   data-background-position: 50% 5%, center
server: shiny
---

```{r, setup, include=FALSE}
library(knitr)
library(shiny)
IRT <- function(theta, a = 1, b = 0, c = 0,e = 1) {
  y <- c + (e - c) * exp(a * (theta - b)) / (1 + exp(a * (theta - b)))
  y[is.na(y)] = 1
  return(y)
}

@@

$$P(x_{pi} = 1|\theta_p, b_i, a_i) = \frac{\exp[a_i(\theta_p - b_i)])}{1 + \exp[a_i(\theta_p - b_i)])}$$

@@

<font size = "5">

$Q_m \subset B = 2^{||B||}-2$ STFs of different lengths considering different item combination can be developed from $B$. That's what Bruto does

1. $\mathbf{TIF}^{Q_m} =  \frac{\sum_{i \in Q_m} IIF_i}{||Q_m||}$
2. $\overline{\Delta}_{\mathbf{TIF}^{Q_m}} =  \mathit{mean}(|\mathbf{TIF}^* - \mathbf{TIF}^{Q_m}|)$

$Q_{bruto} = \arg \min_{Q_m \subset B} \overline{\Delta}_{\mathbf{TIF}^{Q_m}}$
</font>

@@

$Q^k \subset B \setminus Q^{k-1}$: Set of item indexes selected for inclusion in the STF up to iteration $k$ ($Q^0 = \emptyset$)

@@

At $k = 0$, $Q^0 = \emptyset$. For $k \geq 1$, iterate

1. $Q^{k+1} = Q^k \cup \arg \min |TIF^* - IIF_{i \in B \setminus Q^k}|$

2. $TIF_{i \in B\setminus Q^{k+1}}^{k+1} = \frac{IIF_{i \in Q^{k+1}} + IIF_{i \in B \setminus Q^{k+1}}}{||Q^{k+1}|| + 1}$

Until the termination criterion $|TIF^* - TIF^{k+1}| \geq |TIF^* - TIF^{k}|$ is true $\rightarrow$ $Q_{frank} = Q^{k+1}$

[^1]: <font size="3"><font size="6">**FRANK**</font>ly I didnt' know how to name it</font>

@@

1. Generate an item bank $B$ of $6$ items: 
			- Difficulty parameters: $\mathcal{U}(-3, 3)$
			- Discrimination parameters:  $\mathcal{U}(.90, 2.0)$

@@

>>>>

---
title: "IRT procedures for shortening tests"
author: "Ottavia M. Epifania & Friends"
format: 
  revealjs: 
    theme: mytheme.scss
    logo: "www/psicostat.jpg"
    footer: "Psicostat meeting"
background-opacity: "0.45"
title-slide-attributes: 
   data-background-image: "https://psicostat.github.io/shine-bright/gbu.jpg"
   data-background-size: 450px, cover
   data-background-position: 50% 15%, center
server: shiny
---

```{css include = FALSE}
.reveal .slide-logo {
  height: 100px !important;
  width: 100px !important;
  max-width: unset !important;
  max-height: unset !important;
}
```

```{r, setup, include=FALSE}
library(knitr)
library(shiny)
library(ggplot2)
library(tidyverse)
IRT <- function(theta, a = 1, b = 0, c = 0,e = 1) {
  y <- c + (e - c) * exp(a * (theta - b)) / (1 + exp(a * (theta - b)))
  y[is.na(y)] = 1
  return(y)
}

@@

Item Response Function:

$$P(x_{pi} = 1|\theta_p, b_i, a_i) = \frac{\exp[a_i(\theta_p - b_i)])}{1 + \exp[a_i(\theta_p - b_i)]}$$
<br>

Item Information Function:

@@

$\forall Q \in\mathcal{Q} = 2^B \setminus \{\emptyset, B\}$,

1. $\mathbf{TIF}^{Q} =  \frac{\sum_{i \in Q} IIF_i}{||Q||}$
2. $\overline{\Delta}_{\mathbf{TIF}^{Q}} =  \mathit{mean}(|\mathbf{TIF}^* - \mathbf{TIF}^{Q}|)$

$Q_{bruto} = \arg \min_{Q \in \mathcal{Q}} \overline{\Delta}_{\mathbf{TIF}^{Q}}$

@@

$Q^k \subset B$: Set of item selected for inclusion in the STF up to iteration $k$ ($Q^0 = \emptyset$)

@@

At $k =0$, $TIF^0 = (0, 0, \ldots, 0)$, $Q^0 = \emptyset$, iterate

1. $A^k = B \setminus Q^k$

2. $\forall i \in A^k$, $PIF_{i}^k = \frac{TIF^k + IIF_{i}}{||Q^k||+1}$

3. $D = \arg \min_{i \in A^k} |TIF^* - PIF_i^k|$

Termination criterion: $|TIF^* - PIF_D^{k}| \geq |TIF^* - TIF^{k-1}|$:

  - If false, $k = k + 1,$ $Q^{k+1} = Q^k \cup \{D\}$, restart from 1

  - If true, stops, $Q_{Frank} = Q^k$

[^1]: <font size="3"><font size="6">**FRANK**</font>ly I didn't know how to name it</font>

@@

100 data frames

1. Generate an item bank $B$ of $6$ items:

      - Difficulty parameters: $\mathcal{U}(-3, 3)$

      - Discrimination parameters:  $\mathcal{U}(.90, 2.0)$

@@

## Results

::: {.panel-tabset}

## So...

```{r}

load("gbu-comparison.RData")
mycomparison = item_comparison[, c("nbruto", "nfrank", "nila")]
mycomparison$iter = 1:100
ic_long = pivot_longer(mycomparison, cols = !iter)

ic = ic_long %>%  
  group_by(name) %>%  
  summarise(min = min(value), max = max(value), mean = mean(value), median = mean(value), 
            sd = sd(value))
ggplot(ic, 
       aes(x = name, y = mean, color = name)) + geom_point(shape = 5, size = 6) + ylim(0,6) + 
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width = .2)+ 
  geom_point(aes(x = name, y = min), shape = 19, size = 6) + geom_point(aes(x = name, y = max), 
                                                                       shape = 15, size = 6) + 
  geom_point(aes(x = name, y = median), shape = 4, size = 6) + 
  scale_color_manual(values = c("royalblue","red",  "seagreen")) + 
  ylab("Number of items") + theme_light() + 
  theme(axis.text = element_text(size = 26),
        axis.title.y = element_text(size = 28),
        axis.title.x = element_blank(), 
        legend.position = "none")

## In summary

```{r}
ggplot(sum_algo, 
       aes(x = l_stf, y = mean, color = algo, 
           shape = algo)) + geom_line(linewidth = 1.2) + geom_point(size = 4) + theme_light() +
  ylab(expression(paste(TIF^"* - ", TIF[STF]))) +  xlab("Length STF") +
  scale_color_manual(values = c("royalblue","red",  "seagreen")) +
  scale_x_continuous(n.breaks = 6)  + theme(axis.title = element_text(size = 28), 
        axis.text = element_text(size = 26), 
        legend.position = c(.7,.9), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 24), 
        legend.key.size = unit(2, "cm")) +
  guides(color=guide_legend(ncol=3))

## Spill the tea

```{r}
ggplot(allalgo, 
       aes(x = factor(l_stf), y = value, color = algo)) + geom_boxplot() + theme_light() +
  ylab(expression(paste(TIF^"* - ", TIF[STF]))) +  xlab("Length STF") +
  scale_color_manual(values = c("royalblue","red",  "seagreen")) +
  theme(axis.title = element_text(size = 28), 
        axis.text = element_text(size = 26), 
        legend.position = c(.7,.9), 
        legend.title = element_blank(), 
        legend.text = element_text(size = 24), 
        legend.key.size = unit(2, "cm")) +
  guides(color=guide_legend(ncol=3))
```

## When $||Q_{ILA}|| = ||Q_{Frank}|| = ||Q_{Bruto}||$
### But $Q_{Ila} \neq Q_{frank} \neq Q_{bruto}$

```{r}
library(patchwork)
all_equals = as.numeric(rownames(select_equal1[select_equal1$is_equal_all == FALSE, ])
)
wrap_plots(g[all_equals])
```

## In the end

<br>

Mathematically, we are at loss

Psychologically? I don't know

Do they work? Hopefully

::: {.callout-tip}
## Acknowfndabfjknc

Pasquale Anselmi, Egidio Robusto, Livio Finos, Gianmarco Altoè

</br>

. . .

::: {.callout-note}

## This happened too

Roses are red, violets are blue,

my computer broke down on these simulations,

and I broke down too
:::
